---
format: revealjs
---

## Today's Agenda {background-image="./Images/background-scales_justice_v3.png" .center}

```{r}
# background-size="1920px 1080px"
library(tidyverse)
library(readxl)

# Dataset 1 - CAT Ratifications: DATA-Ratifications-CAT-2026-07.xlsx
rats <- read_excel("../../Data/DATA-Ratifications-CAT-2026-07.xlsx", na = "NA")
    
# Dataset 2 - CAT Ratifications + CIRIGHTS data: DATA-CAT_CIRIGHTS-1981-2022.xlsx
ciri <- read_excel("../../Data/DATA-CAT_CIRIGHTS-1981-2022.xlsx", na = "NA")
    
# Dataset 3 - CAT Ratifications + V-Dem data: DATA-CAT_VDem-1980-2025.xlsx
vdem <- read_excel("../../Data/DATA-CAT_VDem-1980-2025.xlsx", na = "NA")
```

<br>

**Section 2: How effective have international institutions been in reducing the use of torture by governments around the world?**

- Simple bivariate data analyses

<br>

::: r-stack
Justin Leinaweaver (Fall 2026)
:::

::: notes

**Prep for Class**

1. Review Canvas submissions

2. Datasets

    - Dataset 1 - CAT Ratifications: DATA-Ratifications-CAT-2026-07.xlsx
    
    - Dataset 2 - CAT Ratifications + CIRIGHTS data: DATA-CAT_CIRIGHTS-1981-2022.xlsx
    
    - Dataset 3 - CAT Ratifications + V-Dem data: DATA-CAT_VDem-1980-2025.xlsx

<br>

**SLIDE**: Today we ...

<br>

**References**

- [UNTC Page](https://treaties.un.org/Pages/ViewDetails.aspx?src=IND&mtdsg_no=IV-9&chapter=4&clang=_en)

- [Treaty Text](https://treaties.un.org/doc/Treaties/1987/06/19870626%2002-38%20AM/Ch_IV_9p.pdf)

:::



## {background-image="./Images/background-scales_justice_v3.png" .center}

Canvas Q1: How long did it take for a majority of the world's countries to ratify the CAT? How long to reach 3/4 of the countries in the dataset?

::: notes

Fall 2020 07-2

- Map of world with all rats by 2017
- Barplot: cumulative ratifications by year
- Barplot: cumulative ratifications by year + slope lines showing steep to start and slow to finish split around 2001
- Map of world with rats by year (yellow to red color showing early vs late)

:::



## {background-image="./Images/background-scales_justice_v3.png" .smaller}

::: {.panel-tabset}

### Map (Current)

```{r, echo=FALSE}

```

### Map (By Year)

```{r, echo=FALSE}

```


### Rats (Cumulative)

```{r, echo=FALSE}

```


### Rats (Slopes)

```{r, echo=FALSE}

```

:::

::: notes

Fall 2020 07-2

- Map of world with all rats by 2017
- Barplot: cumulative ratifications by year
- Barplot: cumulative ratifications by year + slope lines showing steep to start and slow to finish split around 2001
- Map of world with rats by year (yellow to red color showing early vs late)

:::






## {background-image="./Images/background-scales_justice_v3.png" .center}

Canvas Q2: According to the CIRIGHTS data project, did countries that ratified the CAT reduce their use of torture in the following years?

::: notes

- Line plot across time: Ratifiers in three groups (1984 torture status 2s vs 1s vs 0s)
- Repeat above but add line of best fit
- Dot plot with line of best fit: Average torture score x years since ratification
- Line plot: Average torture score ratifiers vs non-ratifiers

:::



## {background-image="./Images/background-scales_justice_v3.png" .center}

Canvas Q3: According to the V-Dem data project, did countries that ratified the CAT reduce their use of torture in the following years?

::: notes



:::





